CF Registry Analysis

### 2024-02-16 Christine Chan/CF Registry/cf_registry_analysis.qmd
forest plot funciton mostly done

@@

```{r}
forest_plot <- function(mod, time_varying = NULL, labels = labels) {
  coeff <- data.frame(summary(mod)$coefficients)
  coeff$ll <- coeff$coef - (1.96 * coeff$se.coef.)
  coeff$ul <- coeff$coef + (1.96 * coeff$se.coef.)
  # Clean up variable names
  l <- unlist(sapply(labels, paste0, ": "))
  coeff$names <- rownames(coeff)
  coeff$names <- str_replace_all(coeff$names, l)
  coeff$names <- ordered(coeff$names, levels = unique(coeff$names))
  # Exponentiate
  coeff$exp.ll <- exp(coeff$ll)
  coeff$exp.ul <- exp(coeff$ul)
  # Plot
  p <- ggplot(coeff, aes(y = fct_rev(names))) +
    geom_point(aes(x = exp.coef.), shape = 15, size = 3) +
    geom_linerange(aes(xmin = exp.ll, xmax = exp.ul)) +
    geom_vline(xintercept = 1, linetype = "dashed") +
    xlab("Hazard Ratio") +
    theme_classic() +
    theme(axis.ticks.y = element_blank(), axis.title.y = element_blank())
}
```

@@

```{r}
# Formula for base model
x <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation", "A_pulmonarycomplications1", "A_corticosteroids1"
)
y <- "Surv(tstart, tstop, cfrd_status_annual)"
f <- as.formula(paste0(y, "~", paste0(x, collapse = "+")))
cox_res <- coxph(f, data = s, cluster = eDWID)
# Plot survival
# ggsurvplot(survfit(cox_res),
#   data = s, risk.table = T, cumevents = T, fun = "event",
#   xlim = c(2003, 2018), break.x.by = 1
# )
# Forest plot and coefficients table
ggforest(cox_res, data = s)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
# ggcoxzph(test.ph)
# ggcoxdiagnostics(cox_res,type = "martingale")
# ggcoxdiagnostics(cox_res,type = "deviance")
# ggcoxdiagnostics(cox_res,type = "score")
# ggcoxdiagnostics(cox_res,type = "schoenfeld")
```

>>>>

# Males to reference group
annual$Gender <- factor(annual$Gender, levels = c("M", "F"))
# Hispanic no as reference
annual$Hispanicrace <- factor(annual$Hispanicrace, levels = c("No", "Yes"))

@@

```{r}
forest_plot <- function(mod, time_varying = NULL, labels = labels) {
  coeff <- tbl_regression(cox_res,
    exponentiate = T,
    pvalue_fun = function(x) x
  ) %>%
    add_n(location = "level") %>%
    as_tibble()
  coeff_n <- tbl_regression(cox_res, exponentiate = T) %>%
    add_n(location = "label") %>%
    as_tibble()
  coeff$N2 <- coeff_n$`**N**`
  # Combine Ns
  coeff <- coeff %>% unite(N, `**N**`, N2, remove = F, na.rm = T)
  # Separate CI column
  coeff$hr <- as.numeric(coeff$`**HR**`)
  coeff$ll <- as.numeric(sapply(str_split(coeff$`**95% CI**`, ", "), "[", 1))
  coeff$ul <- as.numeric(sapply(str_split(coeff$`**95% CI**`, ", "), "[", 2))
  # Add N to names
  coeff$names <- paste0(coeff$`**Characteristic**`, " (N = ", coeff$N, ")")
  coeff$names <- ordered(coeff$names,
    levels = unique(coeff$names)
  )
  # P values
  coeff$sig <- coeff$`**p-value**` < 0.05
  coeff$p <- style_pvalue(coeff$`**p-value**`)
  # Row types
  coeff$type <- "level"
  coeff$type[!is.na(coeff$N2)] <- "variable"
  coeff$type[!is.na(coeff$`**N**`) & is.na(coeff$`**HR**`)] <- "reference"
  # HR label
  coeff$hr_lab <- paste0(coeff$`**HR**`, " (", coeff$`**95% CI**`, ")")
  coeff$hr_lab[coeff$hr_lab == "NA (NA)"] <- NA
  coeff$hr_lab[coeff$type == "reference"] <- "Reference"
  # Plot
  p_left <- ggplot(coeff, aes(y = fct_rev(names))) +
    geom_text(aes(x = 0, label = names),
      hjust = 0,
      fontface = ifelse(coeff$type == "variable", "bold", "plain")
    ) +
    geom_text(aes(x = 1, label = hr_lab), hjust = 0) +
    theme_void() +
    coord_cartesian(xlim = c(0, 2)) +
    geom_stripes()
  p <- ggplot(coeff, aes(y = fct_rev(names))) +
    geom_point(aes(x = hr), shape = 15, size = 3) +
    geom_linerange(aes(xmin = ll, xmax = ul)) +
    geom_vline(xintercept = 1, linetype = "dashed") +
    xlab("Hazard Ratio") +
    theme_classic() +
    theme(
      axis.line.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.line.x = element_blank()
    ) +
    geom_stripes()
  # p values
  p_right <- ggplot(coeff, aes(y = fct_rev(names))) +
    geom_text(aes(x = 0, label = p),
      hjust = 0,
      fontface = ifelse(coeff$sig, "bold", "plain")
    ) +
    theme_void() +
    geom_stripes()
  # Layout
  p_left + p + p_right +
    plot_layout(widths = c(1, 2, 0.25), guides = "collect") &
    theme(plot.margin = margin(0, 0, 0, 0, "pt"))
  # Return each plot separately - will need to be combined outside of the function
  return(c(p_left, p, p_right))
}
```

@@

```{r}
# Formula for base model
x <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation", "A_pulmonarycomplications1", "A_corticosteroids1"
)
y <- "Surv(tstart, tstop, cfrd_status_annual)"
f <- as.formula(paste0(y, "~", paste0(x, collapse = "+")))
cox_res <- coxph(f, data = s, cluster = eDWID)
# Plot survival
# ggsurvplot(survfit(cox_res),
#   data = s, risk.table = T, cumevents = T, fun = "event",
#   xlim = c(2003, 2018), break.x.by = 1
# )
# Forest plot and coefficients table
forest_plot(cox_res)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
# ggcoxzph(test.ph)
# ggcoxdiagnostics(cox_res,type = "martingale")
# ggcoxdiagnostics(cox_res,type = "deviance")
# ggcoxdiagnostics(cox_res,type = "score")
# ggcoxdiagnostics(cox_res,type = "schoenfeld")
```
